Constraining the High Density Nuclear Symmetry Energy with Pions

### 2020-06-24 chapters/ch_intro.tex
Bill edits added Experiment section

It has been shown that most of the $\Delta$'s are produced in the early dense regions of the collision \cite{mingzhang}. Figure~\ref{fig:deltaProduction} shows the average local density (c) which $\Delta$'s are produced and the number in the system (b), as a function of time in the simulation of Au + Au collisions at \SI{400}{\MeVA}. Panel (a) shows the density distribution of the density at the moment of creation for $\Delta$'s. Since the average lifetime of the $\Delta$  is $\tau_{\Delta} = \SI{1.7}{\femto\metre\per\clight}$, the $\Delta$ resonance has very little time to be affected by the medium before decaying into a $\pi$ and nucleon. Thus the outgoing $\pi$ is contains information on the high density region of the collision.

@@

The branching ratio of the various flavors of $\Delta$'s is given by the Clebsh-Gordon coefficients as shown in \ref{appen:deltadecay}. Here we see that in general proton-proton collisions give rise primarily to  $\pi^+$ and neutron-neutron collisions give rise to primarily $\pi^-$. In this $\Delta$ resonance model the charged pion ratio can be described as,

\begin{equation}
\frac{\pi^-}{\pi^+} = \frac{ 5pp + pn }{5nn + pn}.
\label{eq:deltaModel}
\end{equation}

In this $\Delta$ resonance model, $\pi^-/\pi^+ \approx (N/Z)^2$ where $N/Z$ is the neutron-proton ratio of the dense central collision where they are produced. Pions can be reabsorbed into a $\Delta$ resonance after colliding with the another nucleon in the backward process of $\Delta \leftrightarrow \pi + N$. This process generally dilutes the pion sensitivity to the high density region, since with each absorption and re-emission changes the pion dynamics or even the charge of the pion reflecting the asymmetry at the point of creation and re-emission. Total pion absorption back into two nucleons requires a three body process, where a pion is absorbed creating a $\Delta$ resonance, then another nucleon must collide with the resonance to create two nucleons. Because of this, the total pion absorption (removing pions from the system) is a smaller effect than the absorption re-emission process. In general the $\pi^-$ and $\pi^+$ are connected with n-n and p-p behavior in the high density, early collision, effectively turning the neutrons  measurement into a charged $\pi^-$ measurement, which is much easier to measure experimentally.

\section{Motivation for Thesis}
In an effort to answer the high density behavior of the symmetry energy, we designed a new detector and a set of experiments of Sn + Sn collisions at \SI{270}{\MeVA}. Here we utilized inverse kinematics where the beam is made of a radio-active neutron rich beam impinging on a stable Sn target. We can probe the isospin asymmetry of the symmetry energy by changing the neutron-proton asymmetry of the incoming beam. Pion production has been studied before in stable beams for beam energies of \SI{400}{\MeVA} and above \cite{fopi}. In these previous experiments, only total pion yields were published and no pion spectra were published, and the efficiency of the detector at detecting low energy pions was difficult to estimate.  Also though the pion production increases with the beam energy, the symmetry energy affects the dynamics of the system less. This is because the time which the two nuclei coalesce in a HIC becomes shorter, and the time in which the symmetry energy can affect nucleons is shortened, thus the effect will becomes less dominant at higher beam energies.  The goals of this Thesis was to design and build a high efficiency detector in order to measure pion and light charged particle spectra resulting from HICs. To do this a new Time Projection Chamber (TPC) was made called the SAMURAI pion Reconstruction Ion Tracker (\spirit) Time Projection Chamber (TPC).

>>>>

It has been shown that most of the $\Delta$'s are produced in the early dense regions of the collision \cite{mingzhang}. Figure~\ref{fig:deltaProduction} shows the average local density (c) which $\Delta$'s are produced and the number in the system (b), as a function of time in the simulation of Au + Au collisions at \SI{400}{\MeVA}. Panel (a) shows the density distribution of the density at the moment of creation for $\Delta$'s. Since the average lifetime of the $\Delta$  is $\tau_{\Delta} = \SI{1.7}{\femto\metre\per\clight}$, the $\Delta$ resonance has very little time to be affected by the medium before decaying into a $\pi$ and nucleon. Thus the outgoing $\pi$ contains information on the high density region of the collision.

@@

The branching ratio of the various flavors of $\Delta$'s are given by the Clebsh-Gordon coefficients as shown in \ref{appen:deltadecay}. Here we see that in general p-p collisions give rise primarily to  $\pi^+$ and n-n collisions give rise to primarily $\pi^-$. In this $\Delta$ resonance model the charged pion ratio can be described as,

\begin{equation}
\frac{\pi^-}{\pi^+} = \frac{ 5N^2 + ZN }{5Z^2 + ZN}.
\label{eq:deltaModel}
\end{equation}

In this $\Delta$ resonance model, $\pi^-/\pi^+ \approx (N/Z)^2$ where $N/Z$ is the neutron-proton ratio of the dense central collision where they are produced. Pions can be reabsorbed into a $\Delta$ resonance after colliding with another nucleon in the backward process of $\Delta \leftrightarrow \pi N$. This process generally dilutes the pion sensitivity to the high density region, since with each absorption and re-emission changes the pion dynamics or even the charge of the pion reflecting the neutron-proton asymmetry at the point of creation and re-emission. Total pion absorption, back into two nucleons, requires a three body process; the pion collides with a nucleon creating a $\Delta$ resonance, then another nucleon must quickly collide with the resonance to create two nucleons. Since this is a three body process,h total pion absorption (removing them from the system) is a smaller effect than the absorption and re-emission process. Aside from these two effects which degrade the observable, in general the $\pi^-$ and $\pi^+$ are connected with n-n and p-p behavior in the high density, early collision; effectively turning the neutron measurement into a charged $\pi^-$ measurement, which is much easier to measure experimentally, and probing the high density region more than any other observable.

\section{Motivation for Thesis}
In an effort to answer the high density behavior of the symmetry energy, we designed a new detector and a set of experiments of Sn + Sn collisions at \SI{270}{\MeVA}. Here we utilized inverse kinematics where the beam is made of a radio-active neutron rich beam impinging on a stable Sn target. We can probe the isospin asymmetry of the symmetry energy by changing the neutron-proton asymmetry of the incoming beam. Pion production has been studied before in stable beams for beam energies of \SI{400}{\MeVA} and above \cite{fopi}. Yet in these previous experiments, only total pion yields were published and no pion spectra was not published. It was also known that the efficiency analysis of the FOPI detector was particularly difficult especially for low energy pions. Although the pion production increases with the beam energy, the effects of the symmetry energy decrease. This is because higher beam velocities mean the time two nuclei spend together becomes shorter, shortening the time in which the symmetry energy can affect nucleons.  The goal of this Thesis was to design and build a high efficiency detector in order to measure pion and light charged particle spectra resulting from HICs. To do this a new Time Projection Chamber (TPC) was made called the SAMURAI pion Reconstruction Ion Tracker (\spirit) Time Projection Chamber (TPC).
